---
title: "Introduction to Dataframes - Lab 2"
output: html_notebook
---

```{r}
library(sparklyr)
spark <- spark_connect(master = "local")

pagecountsAllDF <- spark_read_parquet(spark, "datasets/wikimedia_edits.parquet")
```

## task

- Filter the result to the en Wikipedia project.
- Find all the articles where the name of the article ends with `Washington` (presumably "George Washington", "Martha Washington, etc)
- Return all records as an array to the Driver.
- Total the requests for the Washingtons and assign it to the variable `totalWashingtons`.
Assign your array of Washingtons (the return value of your action) to the variable `washingtons`.

Bonus: repeat the exercise for the Adams

- Find all the articles where the name of the article ends with `Adams` (presumably "John Adams", "John Quincy Adams", etc)
- Return all records as an array to the Driver.
- Total the requests for the Adamas and assign it to the variable `totalAdams`.
- Assign your array of Adamas (the return value of your action) to the variable `adams`.

>But you cannot do it the same way twice:
>In the filter, don't use the same conditional method as the one used for the >Washingtons.
>Don't use the same action as used for the Washingtons.

```{r}
# TODO fill in
```


## cleanup

Finally, close the spark session again.
```{r}
spark_disconnect(spark)
```